---
title: "Assignment 9"
author: "Jiho Kim"
date: "10/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

To choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.


```{r}
library(jsonlite)
library(dplyr)
```

## Connecting to the API

I set up account to retreive appropriate API key and I used NY Times Movie review API and used my API key and URL to retreive the data. 
```{r}
apikey <- 'IBAZ6dq08t5OBbTyp9l2DNvm44OhA9DT'
http <- "https://api.nytimes.com/svc/movies/v2/reviews/search.json?api-key=IBAZ6dq08t5OBbTyp9l2DNvm44OhA9DT"
```

## Read JSON DATA

Reading content in JSON 
```{r}
moviedf <- fromJSON(http)
View(moviedf)
class(moviedf)
```

## Transforming into Data frame

Movie data is displayed in list so I have to convert the list to a dataframe with only variabloes of interest (title, rating, critics pick, and opening date ).

```{r}
finalmovie <- cbind(moviedf$results$display_title, moviedf$results$mpaa_rating,
                    moviedf$results$critics_pick,moviedf$results$opening_date)
finalmovie <- as.data.frame(finalmovie)
View(finalmovie)
## Renaming the column names

colnames(finalmovie) <- c('Title','Rating','Critics_Pick','Opening_date')

finalmovie
```




