---
title: "Project 2"
author: "Jiho Kim"
date: "10/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data 1

The original data set was obtained from Kaggle and cosists of student performance and the data was recreated as a wide format for the assignment. I want to see if the education level of parents affect the perfomrane of their children.

## Set up

```{r}
library(dplyr)
library(tidyverse)
library(readr)
library(tidyr)
```

## Data

```{r}
urlfile <- 'https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/project%202/Student%20Performace%20.csv'
mydata <- read.csv(url(urlfile))
mydata <- as.data.frame(mydata)

view(mydata)

```

## Tidying

So, when we look at the data we don't need the first column and tyding and transforming needs to be done.

```{r}
mydata <- mydata[-1]
```
```{r}
newdata <- gather(mydata, "Student", "Score", 2:21, factor_key=TRUE)

newdata <- spread(newdata, "X1", "Score")

newdata <- newdata[,c(1,3,2,4,5)]



```

```{r}
colnames(newdata) <- c('student', 'education.level.parents.','math.score','reading.score','writing.score')
```

## Transform Data

```{r}
high_ed <- newdata %>%
  filter(education.level.parents. == "master's degree"| education.level.parents. == "bachelor's degree")

mid <- newdata %>%
  filter(education.level.parents. == 'some college' | education.level.parents. == "associate's degree")

low <- newdata %>%
  filter(education.level.parents. == "high school")
```

## Analysis

```{r}
u_avg_math <- sum(as.numeric(high_ed$`math.score`)) / nrow(high_ed)
u_avg_read <- sum(as.numeric(high_ed$`reading.score`)) / nrow(high_ed)
u_avg_write <- sum(as.numeric(high_ed$`writing.score`)) / nrow(high_ed)

m_avg_math <- sum(as.numeric(mid$`math.score`)) / nrow(mid)
m_avg_read <- sum(as.numeric(mid$`reading.score`)) / nrow(mid)
m_avg_write <- sum(as.numeric(mid$`writing.score`)) / nrow(mid)

l_avg_math <- sum(as.numeric(low$`math.score`)) / nrow(low)
l_avg_read <- sum(as.numeric(low$`reading.score`)) / nrow(low)
l_avg_write <- sum(as.numeric(low$`writing.score`)) / nrow(low)
```

## Bar Grpah
```{r}
barplot(c(u_avg_math, m_avg_math, l_avg_math), beside=TRUE, main = "Avg Math Score v Parent Education Level", xlab = "Parent Education Level", ylab = "Avg Math Score", ylim = c(0, 70), names.arg = c("Upper", "Mid", "Lower"))
```

```{r}

barplot(c(u_avg_read, m_avg_read, l_avg_read), beside=TRUE, main = "Avg Reading Score v Parent Education Level", xlab = "Parent Education Level", ylab = "Avg Reading Score", ylim = c(0, 75), names.arg = c("Upper", "Mid", "Lower"))
```

```{r}
barplot(c(u_avg_write, m_avg_write, l_avg_write), beside=TRUE, main = "Avg Writing Score v Parent Education Level", xlab = "Parent Education Level", ylab = "Avg Writing Score", ylim = c(0, 75), names.arg = c("Upper", "Mid", "Lower"))
```

## Conclusion

Unlinke what I expected, lower group performed the best out of three groups. 


## DATA 2

The data is from Kaggle and contains coaches of 2020 olympics. The data was recreated for this assignment to test which discline of sports had the most coaches in 2020 olympics.

```{r}
url2 <- 'https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/project%202/Coachescsv.csv'

olympics <- read.csv(url2)
as.data.frame(olympics)


```

## Tidying

Looking at the data, the datatable should be transposed. And first 2 rows and last columns are not neccessary.

```{r}
newolym<-t(olympics)

newolym <- newolym[-c(1:2),-4]


```

## Transforming
```{r}
newolym<-as.data.frame(newolym)
colnames(newolym) <- c('name','country','discipline')
rownames(newolym)<-c(1:nrow(newolym))
final<-newolym%>%
  group_by(discipline)%>%
  summarise(count = n())
 
```


## Analysis

```{r}
ggplot(final,aes(x=discipline, y=count))+geom_bar(stat='identity')
```

According to the plot, artistic swimming had the most coaches and water polo had the least coaches.


## DATA 3

The data is from World Bank and contains countries gdp growth and we don't need first 2 columns. I want to compare GDP growth rate of North Korea and South Korea.

```{r}
url3<- 'https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/project%202/f42e96ef-91b8-47d6-94c9-769c6bc8de40_Data.csv'
gdp<-read.csv(url3)
gdp<-gdp[218:434,-c(1:2)]
```

## Tydying and Cleaning

I want to comapre gdp growht rate of North Korea and South Korea since 2000.

```{r}
gdpnew<-gdp %>%
  pivot_longer(cols=3:ncol(gdp),names_to = 'year',values_to = 'gdp.growth')
gdpfinal <- gdpnew%>%filter(Country.Code == 'PRK' | Country.Code == 'KOR' )
gdpfinal<-mutate(gdpfinal, Year = rep(seq(2000:2015),2))
gdpfinal$gdp.growth<-round(as.numeric(gdpfinal$gdp.growth),2)

```

## Analysis

```{r}
ggplot(gdpfinal,aes(x=Year, y = gdp.growth))+ggtitle('GDP growth since 2000')+geom_line(aes(color=Country.Name))
```

South Korea expreienced spike of increase in GDP in 2006 while North Korea's GDP decreased from 2004. 


