---
title: "Assignment 3"
author: "Jiho Kim"
date: "9/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Provide code that identifies the major that contain either "DATA" or " Statistics" in dataset.

```{r data}
majors <- read.csv('https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/all-ages.csv', header = T)
majordf <- data.frame(majors)
data_stat<-subset(majordf, grepl("DATA",Major) | grepl("STATISTICS", Major), Major)
data_stat
```


# 2. Write code that transforms the data 

```{r fruitlist}
fruitlist <- c('bell pepper','bilberry', 'blackberry', 'blood orange','blueberry','cantaloupe', 'chili pepper','cloudberry','elderberry','lime','lychee','mulberry','olive','salal berry')

dput(as.character(fruitlist))
```

# 3 Decribe, in words, what these expression will match

(.)\1\1

This expression would match character that appears 3 times in a row.

(.)(.)\2\1

matches a letter in order of 1st letter-2nd letter-1st letter-2nd letter.

(..)\1

It would match any 2 character repeated consecutively. 

(.).\1.\1

It would match with first character repeated 3 times with second character repeated 2 times in between

(.)(.)(.).*\\3\\2\\1

It would match with any 3 characters followed by any characters ending with the original 3 characters in reverse


# 4 Construct regular expressions

```{r word list}
library(tidyverse)
words <- c('mom','dead','test','banana','church','eleven','Mississippi')
```

start and end with the same character:
```{r same character}
first_end <- str_subset(words, "^(.).*\\1$")
first_end
```

contain a repeated pair of letters:
```{r repeat}
repeatword <- str_subset(words, "([A-Za-z][A-Za-z]).*\\1")
repeatword
```

conatin one letter repeated in at least three places:
```{r threetimes}
threetimes <- str_subset(words, "([A-Za-z]).*\\1.*\\1")
threetimes
```












