---
title: "Assignment 5"
author: "Jiho Kim"
date: "9/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reaading the Data

```{r data}
library(readr)
library(tidyverse)
urlfile <- 'https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/Assignment%205/israeli_vaccination_data_analysis_start.csv'
mydata <- read.csv(url(urlfile), header = TRUE)
data.frame(mydata)

```

## Tidy Data

```{r}

colnames(mydata) <- c('Age', 'Unvaccinated', 'Vaccinated', 'Severe_Unvaccinated_Per100K', 'Severe_Vaccinated_Per100K')

mydata <- mydata[2:5,1:5]

```

```{r}
pop_underfifty <- mydata[2,2:3]
pop_underfifty <- pop_underfifty %>% rename(percent_unvaccinated = Unvaccinated, percent_vaccinated = Vaccinated)
```

```{r}
pop_overfifty <- mydata[4,2:3]
pop_overfifty <- pop_overfifty %>% rename(percent_unvaccinated = Unvaccinated, percent_vaccianted = Vaccinated)
```

```{r}
mydata <- add_column(mydata, pop_underfifty, .after = 'Vaccinated')
```

```{r}
mydata[4:5,3:4] <- pop_overfifty
```

```{r}
mydata <- mydata[-c(2,4,5),]

```

```{r}
mydata[2,4:5] <- pop_overfifty
```

```{r}
mydata
```

## Question 1

Yes we do have enough information to caclculate the total population. Total population would represent poeple who are full vaccinated, not fully vaccinated or un-Vaccinated. 

## Question 2

Efficacy = 1-(% of fully vaxed severe cases) / (% not vaxed severe cases per 100k)

```{r}
class(mydata$Unvaccinated)='Numeric'
class(mydata$Vaccinated)='Numeric'
class(mydata$percent_unvaccinated)='Numeric'
class(mydata$percent_vaccinated)= 'Numeric'
class(mydata$Severe_Unvaccinated_Per100K)='Numeric'
class(mydata$Severe_Vaccinated_Per100K)='Numeric'
```

For Some reasons, I kept getting error when trying to calcualte the efficacy and NA was introduced continously.

