---
title: "Assignment 10"
author: "Jiho Kim"
date: "10/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In Text Mining with R, Chapter 2 looks at Sentiment Analysis. In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document. You should provide a citation to this base code. You’re then asked to extend the code in two ways:

  Work with a different corpus of your choosing, and
  Incorporate at least one additional sentiment lexicon (possibly from another R package that    you’ve found through research).

## Set Up

```{r}
library(tidytext)
library(tidyr)
library(dplyr)
library(janeaustenr)
library(stringr)
library(ggplot2)
library(wordcloud)
library(reshape2)
```

## Example Code

The following code was retreived directly from : 

Silge, J., &amp; Robinson, D. (2017). Text mining with R: A tidy approach. O'Reilly Media.

## 2.1 The Sentiments datasets

```{r}
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

## 2.2 Sentiment analysis with inner join

```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## 2.3 Comparing 3 sentiment dictionaries

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(pride_prejudice %>% 
                            inner_join(get_sentiments("bing")) %>%
                            mutate(method = "Bing et al."),
                          pride_prejudice %>% 
                            inner_join(get_sentiments("nrc") %>% 
                                         filter(sentiment %in% c("positive", 
                                                                 "negative"))) %>%
                            mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}

get_sentiments("nrc") %>% 
     filter(sentiment %in% c("positive", 
                             "negative")) %>% 
  count(sentiment)
```

```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```

## 2.4 Most common positive and negative words

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts
```

```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

## 2.5 Wordclouds
```{r}

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

```{r}

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

## 2.6 Looking at units beyond just words

```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
```

```{r}
p_and_p_sentences$sentence[2]
```
```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

## New Corpus

I will be using Adventures of Huckleberry Finn data from gutenberg packages.

```{r}
library(gutenbergr)

# get ID and Downlaod the data 
gutenberg_metadata %>%
  filter(title == 'Adventures of Huckleberry Finn')

HuckFinn <- gutenberg_download(76)
HuckFinn
```

## Tidying

```{r}
HuckFinn <- HuckFinn[c(515:nrow(HuckFinn)),]
HuckFinn <- HuckFinn %>%
  select(-gutenberg_id)%>%
  filter(text != "")%>%
  mutate(author = 'Mark Twain',
         line = row_number(),
         chapter = cumsum(str_detect(text, regex("CHAPTER [\\dIVXLC]", ignore_case =  TRUE))))
#identify tokens and remove stop words
HuckFinnTidy <- HuckFinn %>%
  unnest_tokens(word,text)%>%
  anti_join(stop_words)
```

## Sentiment Analysis

I will be using Loughran lexicon to perform sentiment analysis

```{r}
get_sentiments('loughran')

HuckFinn.lexi<- HuckFinnTidy %>%
inner_join(get_sentiments("loughran")) %>%
count(word, sentiment, sort = TRUE) %>%
ungroup()
```

## Top 10 sentiments

```{r}
HuckFinn.lexi %>%
group_by(sentiment) %>%
top_n(10) %>%
ungroup() %>%mutate(word = reorder(word, n)) 
```

## positive and negative sentiments for nrc
```{r}
HuckFinn.nrc <- HuckFinnTidy %>%
inner_join(get_sentiments("nrc")) %>%
count(word, sentiment, sort = TRUE) %>%
ungroup()
```

## Plot for nrc
```{r}
huckfinnposnegnrc<-HuckFinn.nrc %>%filter(sentiment =='positive' | sentiment =='negative')
huckfinnposnegnrc %>%
     filter(n > 20) %>%
     mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
     mutate(word = reorder(word, n)) %>%
     ggplot(aes(word, n, fill = sentiment))+
     geom_col() +
     coord_flip() +
     labs(y = "Count")
```


## Positive and Negative sentiments for Loughran

```{r}
Positive.Negative<-HuckFinn.lexi %>%filter(sentiment =='positive' | sentiment =='negative')
```

## Plot for positive and negative sentiments for Loughran
```{r}
Positive.Negative %>%
     filter(n > 20) %>%
     mutate(n = ifelse(sentiment == "negative", -n, n)) %>%
     mutate(word = reorder(word, n)) %>%
     ggplot(aes(word, n, fill = sentiment))+
     geom_col() +
     coord_flip() +
     labs(y = "Count")
```

## Word Cloud 

```{r}
HuckFinnTidy %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))
```

## Word Cloud for loughran

```{r}
 HuckFinnTidy %>%
  inner_join(get_sentiments("loughran")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

## Conclusion

Sentiment Analysis is a great tool to understand how narrative and emotions are expressed in certain texts. For this assignment, we analyzed Adventure of Huckleberry Finn using loughran and nrc lexicons. 

