---
title: "Project 1"
author: "Jiho Kim"
date: "9/16/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Descrition

In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents For the first player, the information would be: Gary Hua, ON, 6.0, 1794, 1605

## Load the Data

Chess data is retrieved from github

```{r chessdata}
chessdata <- read.table(url("https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/Project%201/tournamentinfo.txt"), sep = ',' , skip = 4)

chessdata
```


```{r whitespace}
i <- seq(0,192,3)
chessdata <- chessdata[-i]

chessdata

```

```{r matrix }
mchessdata <- matrix(unlist(chessdata), byrow= T)

m1 <- mchessdata[seq(1,length(mchessdata),3)]
m1

m2 <- mchessdata[seq(2,length(mchessdata),3)]
m2
```

```{r capture}
library(stringr)
Id <- as.numeric(str_extract(m1, '\\d+'))
Name <- str_extract(m1, '[A-z].{1,32}')
Name <- str_trim(str_extract(Name, '.+\\s{2,}'))
State <- str_extract(m2, '[A-z]{2}')
Total.points <- as.numeric(str_extract(m1,'\\d+\\.\\d'))
PreRating <- str_extract(m2, 'R:.{8,}-')
PreRating <- as.numeric(str_extract(PreRating, '\\d{1,4}'))
Rounds <- str_extract_all(m1, '[A-Z]\\s{2,}\\d+')
Rounds <- str_extract_all(Rounds, '\\d+')
```

```{r avg}
AvgOpp <- c()

for(i in c(1:length(Rounds))){
  AvgOpp[i] <- round(mean(PreRating[as.numeric(Rounds[[i]])]),0)
}
AvgOpp
```

```{r final}
finalresult <- data.frame(Id,Name,State,Total.points,PreRating,AvgOpp)

finalresult
```

```{r csv}
write.csv(finalresult, 'project1' , append = F)
```

