---
title: "Week 7 Assignment"
author: "Jiho Kim"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

For this Assignment I chose 3 of my favorite books and listed below:

Great Gatsby By F.Scott Fitzerald
1984 by George Orwell
Fight Club by Chuck Palahniuk
The Tailsman by Stephen King and Peter Straub

## Importing HTML 

The HTML is created and uploaded to github.
```{r}
library(RCurl)
library(XML)

html <- getURLContent('https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/assignment%206/book.html')
df.html <- readHTMLTable(html, as.data.frame = TRUE)
df.html
```

## Importing XML
```{r}
xml <- getURLContent('https://raw.githubusercontent.com/jihokim97/CUNY-SPS-MSDS-DATA-607/main/assignment%206/book.xml')
df.xml <- xmlToDataFrame(xml)
df.xml
```
 
## Importing JSON
```{r}
library(jsonlite)
json <- getURLContent('https://github.com/jihokim97/CUNY-SPS-MSDS-DATA-607/blob/main/assignment%206/book.json')
jsondata <- fromJSON(json)
df.json <- as.data.frame(jsondata)
df.json
```

## Compare and Conlusion

The html and xml data frames are identical. I had a trouble writing and reading json file, and I would like a kind feedback on what went wrong. Each structures had its own pros and cons. I prefered xml and html as json was more difficult to create and read.